---
title: "Group project"
author: "Jakub Grad"
date: "2024-11-27"
output:
  pdf_document: 
    latex_engine: lualatex
  html_document: default
  bookdown::pdf_book:
    latex_engine: lualatex
header-includes:
  - \usepackage{amsmath}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
train <- read.csv(file = "kaggle/train.csv")
test <- read.csv(file = "kaggle/test.csv")

# Select only numeric columns from 'train'
train_numeric <- train[, sapply(train, is.numeric)]
test_numeric <- test[, sapply(test, is.numeric)]

linear_regressor <- glm(formula = log_pSat_Pa ~ ., data=train)
test$TARGET <- predict(linear_regressor, newdata=test)

submission_df <- test[, c("ID", "TARGET")]
write.csv(submission_df, "kaggle_submission.csv", row.names = FALSE)
```

```{r, cache=TRUE}
pairs(train[, sapply(train, is.numeric)])
```

```{r, cache=TRUE}
# Compute the correlation matrix
cor_matrix <- cor(train_numeric)

# Create a heatmap of the correlation matrix
heatmap(cor_matrix)
```

Ideas:
Use cross validation before submitting?

Perform Lasso?
```{r, cache=TRUE}
library(glmnet)
any(is.na(train))

x <- model.matrix(log_pSat_Pa ~ ., train_numeric)[,-1]
y <- train_numeric$log_pSat_Pa

grid <- 10^ seq (10,-2, length = 100)

lasso.mod <- glmnet(x, y, alpha = 1, lambda = grid)
plot(lasso.mod)
```
Let's find out the CV error
```{r}
set.seed (17)
cv.error.100 <- rep(0, 100)

cv.lasso <- cv.glmnet(x, y, alpha = 1, lambda = grid, nfolds = 10)

best.lambda <- cv.lasso$lambda.min
best.lambda
best.coefficients <- coef(cv.lasso, s = best.lambda)
```
I've created a model using best coefficients from lasso and it has a lower score than without using lasso. The best lambda returned is the lowest one, implying that lasso is not the right approach. 
```{r}
test$TARGET <- predict(cv.lasso, newx = model.matrix(~ ., test_numeric)[,-1], s = best.lambda)

submission_df <- test[, c("ID", "TARGET")]
write.csv(submission_df, "lasso.csv", row.names = FALSE)
```
Mb i should plot plogsat against different other variables, see what kind of relations are there...
random forest regression, use it, it's apparently good

subset selection: best subset selection, stepwise, ridge regression. reetta suggests PCA, chapter 121

model selection

What about ridge regression?
```{r}
ridge.mod <- glmnet(x, y, alpha = 0, lambda = grid)

plot(ridge.mod, xvar = "lambda", label = TRUE)
```

Cross-Validation for Ridge
```{r}
set.seed(17)

# Perform cross-validation
cv.ridge <- cv.glmnet(x, y, alpha = 0, lambda = grid, nfolds = 10)

# Optimal lambda
best.lambda <- cv.ridge$lambda.min
cat("Best lambda:", best.lambda, "\n")
## Best lambda: 0.01
# Plot cross-validation curve
plot(cv.ridge)

```

submission. Result is 0.7101 which is still worse than regular regression.
```{r}
test$TARGET <- predict(cv.ridge, newx = model.matrix(~ ., test_numeric)[,-1], s = best.lambda)

submission_df <- test[, c("ID", "TARGET")]
write.csv(submission_df, "ridge_submission.csv", row.names = FALSE)
```

Plot plots to show correlation
```{r}
library(ggplot2)

# Subset 1: Select a few numeric predictors
subset_1 <- train_numeric[, c("log_pSat_Pa", "MW", "NumOfAtoms", "NumOfC", "NumOfO")]
# Create pairwise plots for subset_1
pairs(subset_1)
```

```{r}
# Subset 2: Select another few predictors, focusing on other numeric and categorical features
subset_2 <- train_numeric[, c("log_pSat_Pa", "NumOfConf", "NumHBondDonors", "C.C..non.aromatic.", "C.C.C.O.in.non.aromatic.ring")]
# Create pairwise plots for subset_2
pairs(subset_2)
```
```{r}
# Subset 3: A set focusing on some chemical features that may affect log_pSat_Pa
subset_3 <- train_numeric[, c("log_pSat_Pa", "hydroxyl..alkyl.", "aldehyde", "ketone", "carboxylic.acid")]
# Create pairwise plots for subset_3
pairs(subset_3)
```
if present, is terribly unlinear. It also seems that there’s lots and lots of responses for a single input for most of the graphs.

GAMs. Dont work because all variables need to have at least 4 distinct values, and this dataset sometimes has 3.
```{r}
# library(gam)
# library(glmnet)
# 
# # Fit a GAM model (log_pSat_Pa ~ s(*) for smooth terms)
# gam_model <- gam(log_pSat_Pa ~ s(MW, 4) + s(NumOfAtoms, 4) + s(NumOfC, 4) + s(NumOfO, 4) + s(NumOfN, 4) + ...
```

Let’s try regression trees is not doing very well by default (0.5852)
```{r}

library(rpart)

# Fit a regression tree model
regression_tree <- rpart(log_pSat_Pa ~ ., data = train_numeric, method = "anova")

# View the model summary
summary(regression_tree)
```

```{r}
# Visualize the tree (optional, but useful for understanding the splits)
plot(regression_tree)
text(regression_tree, use.n = TRUE, cex = 0.7)
```

```{r}
# Make predictions on the test data
test$TARGET <- predict(regression_tree, newdata = test_numeric)

# Create the submission dataframe
submission_df <- test[, c("ID", "TARGET")]

# Write the submission to a CSV file
write.csv(submission_df, "regression_tree_submission.csv", row.names = FALSE)
```

Let’s try a bigger regression tree:
```{r}
library(tree)

# Fit the regression tree with a bigger tree by passing control
regression_tree_big <- tree(log_pSat_Pa ~ ., data = train_numeric, control = tree.control(nobs = nrow(train_numeric), mindev = 0.0005))

# View the model summary
summary(regression_tree_big)
```

```{r}
# Visualize the tree (optional)
plot(regression_tree_big)
text(regression_tree_big, use.n = TRUE, cex = 0.7)
```

```{r}
# Make predictions on the test data
test$TARGET <- predict(regression_tree_big, newdata = test_numeric)

# Create the submission dataframe
submission_df <- test[, c("ID", "TARGET")]

# Write the submission to a CSV file
write.csv(submission_df, "regression_tree_big_submission.csv", row.names = FALSE)
```

Prune the big tree into the best tree according to CV:
```{r}
# Perform cross-validation to find the best pruned tree size
cv.tree_result <- cv.tree(regression_tree_big)

# Plot the cross-validation results to inspect the best tree size
plot(cv.tree_result$size, cv.tree_result$dev, type = "b", xlab = "Tree Size (Terminal Nodes)", ylab = "Deviance")
```

```{r}
# Find the optimal size (best pruning)
best_size <- cv.tree_result$size[which.min(cv.tree_result$dev)]
best_size <- 10
best_pruned_tree <- prune.tree(regression_tree_big, best = best_size)

# Visualize the best pruned tree
plot(best_pruned_tree)
text(best_pruned_tree, pretty = 0)

```

```{r}
# Make predictions on the test data
test$TARGET <- predict(best_pruned_tree, newdata = test_numeric)

# Create the submission dataframe
submission_df <- test[, c("ID", "TARGET")]

# Write the submission to a CSV file
write.csv(submission_df, "tree_with_10_nodes_submission.csv", row.names = FALSE)

```

```{r}
library(randomForest)
```

```{r}
# Fit a Random Forest model
set.seed(42)  # Set a seed for reproducibility
rf_model <- randomForest(log_pSat_Pa ~ ., data = train_numeric, ntree = 50, importance = TRUE)

# Print the model summary
print(rf_model)
```

```{r}
# View variable importance (optional)
importance(rf_model)

```

```{r}
varImpPlot(rf_model)
```

```{r}
# Make predictions on the test data
test$TARGET <- predict(rf_model, newdata = test_numeric)

# Create the submission dataframe
submission_df <- test[, c("ID", "TARGET")]

# Write the submission to a CSV file
write.csv(submission_df, "random_forest_submission.csv", row.names = FALSE)
```

